VM-Centric Snapshot Deduplication for Converged Cloud Architectures

### 2014-07-19 snapshot_system/paper_arch.tex
re

\subsection{Per Node Software Components} 
%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
    \centering
    %\includegraphics[width=2.75in]{images/socc_arch_cluster}
    \includegraphics[width=3in]{images/socc_arch_cluster}
%\vspace{1em}
    \caption{Data flow during snapshot backup}
    \label{fig:arch_vm}
\end{figure}
As depicted in Figure~\ref{fig:arch_vm}, 
there are four key service components running on each cluster
node  for supporting backup and deduplication: 
1) a virtual block device driver, 2) a snapshot deduplication agent,
3) a snapshot store client to store  and access snapshot data,
and 4)  a PDS client to support PDS metadata access.

@@

%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
  \centering
%  \includegraphics[width=2.75in]{images/sstore_arch}
  \includegraphics[width=3in]{images/sstore_arch}
%\vspace{1em}
  \centering
  \caption{VM snapshot store data structures}
  \label{fig:as_arch}
\end{figure}

@@

For non PDS data, the snapshot store of a VM is  divided into a set of containers and 
each container is approximately 1GB.  The reason for this  divison is to simplify the compaction process
conducted periodically. When  chunks are deleted from old snapshots,
chunks without any reference from other snapshots can be removed by this compaction process.
By limiting the size of a container, we can control the length of each round of compaction.
The compaction  can work on one container at a time and move the in-use data chunks to another container.

Each non-PDS data container is further divided into a set of chunk data groups. Each chunk group is composed of
a set of data chunks and is the basic unit in data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and stores the entire
group as a unit after compression. This  compression can reduce data by several times  in our tested data.
When accessing a particular chunk, its chunk group is retrieved from the storage
and decompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Guo2011,Rhea2008},
retrieval of  a data chunk  group naturally works well with prefetching. 
A  typical chunk group contains 1000 chunks in our experiment.
% with an average size of 200-600 chunks.

%\item
Each non-PDS data container is represented by three files in the DFS:
1) the container data file holds the actual content, 
2) the container index file is responsible for translating a data reference
into its location within a container, and 
3) a chunk deletion log file records all the deletion requests within  the container.

>>>>

\subsection{Per Node Software Components} 
%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
    \centering
    %\includegraphics[width=2.75in]{images/socc_arch_cluster}
    \includegraphics[width=2.8in]{images/socc_arch_cluster}
%\vspace{1em}
    \caption{Data flow during snapshot backup}
    \label{fig:arch_vm}
\end{figure}
As depicted in Figure~\ref{fig:arch_vm}, 
there are four key service components running on each cluster
node  for supporting backup and deduplication: 
1) a virtual block device driver, 2) a snapshot deduplication agent,
3) a snapshot store client to store  and access snapshot data,
and 4)  a PDS client to support PDS metadata access.

@@

%\begin{figure}[htbp]
\begin{figure}[htbp]
%\vspace{1em}
  \centering
%  \includegraphics[width=2.75in]{images/sstore_arch}
  \includegraphics[width=2.8in]{images/sstore_arch}
%\vspace{1em}
  \centering
  \caption{VM snapshot store data structures}
  \label{fig:as_arch}
\end{figure}

@@

For non PDS data, the snapshot store of a VM is  divided into a set of containers and 
each container is approximately 1GB.  The reason for this  division is to simplify the compaction process
conducted periodically. When  chunks are deleted from old snapshots,
chunks without any reference from other snapshots can be removed by this compaction process.
By limiting the size of a container, we can control the length of each round of compaction.
The compaction  can work on one container at a time and move the in-use data chunks to another container.

Each non-PDS container is further divided into a set of chunk data groups. Each chunk group is composed of
a set of data chunks and is the basic unit in data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and stores the entire
group as a unit after compression. This  compression can reduce data by several times  in our tested data.
When accessing a particular chunk, its chunk group is retrieved from the storage
and decompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Guo2011,Rhea2008},
retrieval of  a data chunk  group naturally works well with prefetching. 
A  typical chunk group contains 1000 chunks in our experiment.
% with an average size of 200-600 chunks.

%\item
Each non-PDS data container is represented by three files:
%in the DFS
1) the container data file holds the actual content, 
2) the container index file is responsible for translating a data reference
into its location within a container, and 
3) a chunk deletion log file records all the deletion requests within  the container.

### 2014-07-19 snapshot_system/paper_background.tex
re
figure update

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
In a converged setting with source-side deduplication, the resources that are used to implement snapshot
backup and deduplication are the same resources that must support cloud-hosted
VMs.  Thus the backup service collocated with 
other cloud services has to minimize its resource impact.  
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

\begin{figure}[htb]
    \centering
    %\includegraphics[width=3in]{images/colocated-arch}
    \includegraphics[width=3.25in]{images/converged_arch}
    \caption{VM snapshot backup running on a converged cloud cluster.}
    \label{fig:collocated}
\end{figure}

@@

%Another disadvantage of cross-VM sharing is that it complicates
Snapshot deletion 
%which is the garbage collection in VM environment
occurs frequently because snapshots expire regularly
and reference counting is required to identify chunks that are no loner used.
%Several well developed garbage collection methods do not apply to the
%converged virtualized architecture due to various environment and resource limitations:
Grouped mark-and-sweep~\cite{Guo2011} is effective for deletion by tracking
the dependencies from data containers to metadata groups, but still it carries
a significant cost in a cloud cluster. This is because any data chunk can still be
used by a large number of VMs, and there are multiple rounds of extensive I/O involved
to scan through the meta data of snapshots to check if a chunk is referenced or not.
As the number of machines increases, the I/O cost becomes more expensive.
% and  we show  such that only a subset of
%metadata needs to be scanned in order to reclaim the garbage in any given data container.
%However, data in VM environment tend to be highly dedupable which means
%such dependency could be extensive, therefore scanning a large subset of metadata still carries a significant cost,
% especially in a distributed setting.
Perfect hashing~\cite{Fabiano2013} provides a memory efficient way to represent a large set of
hashes in a compact form, however, it still requires the scanning of all data hashes before any 
garbage collection could begin, which limits its application to an offline batch process. 
In addition, it requires a centralized location
to store the hashing result which is still too big to hold in a converged architecture.

>>>>

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host. 
%The secondary storage can be cluster-based and implemented with a distributed file system~\cite{googlefs03}.   
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
In a converged setting with source-side deduplication, the resources that are used to implement snapshot
backup and deduplication are the same resources that must support cloud-hosted
VMs.  Thus the backup service collocated with 
other cloud services has to minimize its resource impact.  
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

\begin{figure}[htb]
    \centering
    %\includegraphics[width=3in]{images/colocated-arch}
    %\includegraphics[width=3in]{images/converged_arch}
    \resizebox{0.85\linewidth}{!}{
        \input{images/converged_arch}
        %\includegraphics{figures/converged_arch}
    }
    \caption{VM snapshot backup running on a converged cloud cluster.}
    \label{fig:collocated}
\end{figure}

@@

%Another disadvantage of cross-VM sharing is that it complicates
Snapshot deletion 
%which is the garbage collection in VM environment
occurs frequently because snapshots expire regularly
and reference counting is required to identify chunks that are no loner used.
%Several well developed garbage collection methods do not apply to the
%converged virtualized architecture due to various environment and resource limitations:
Grouped mark-and-sweep~\cite{Guo2011} is effective for deletion by tracking
the dependencies from data containers to metadata groups, but still it carries
a significant cost in a cloud cluster. This is because any data chunk can still be
used by a large number of VMs, and there are multiple rounds of extensive I/O involved
to scan through the meta data of snapshots to check if a chunk is referenced or not.
As the number of machines increases, the I/O cost becomes more expensive.
% and  we show  such that only a subset of
%metadata needs to be scanned in order to reclaim the garbage in any given data container.
%However, data in VM environment tend to be highly dedupable which means
%such dependency could be extensive, therefore scanning a large subset of metadata still carries a significant cost,
% especially in a distributed setting.
Perfect hashing~\cite{Fabiano2013} provides a memory efficient way to represent a large set of
hashes in a compact form, however, it still requires the scanning of all data hashes before any 
garbage collection could begin, which limits its application to an offline batch process. 
In addition, it requires a centralized location
to store the hashing result which is still too big to be held in a converged architecture.

Our previous work on parallel deduplication~\cite{zhangusenix13} is a low-cost
solution, but it is not an inline solution and th backup for all VMs needs to be conducted
around the same time. It does not address  snapshot deletion and file system block 
packaging for fault tolerance.

### 2014-07-19 snapshot_system/paper_eval.tex
re

In summary, our experiments show that
version detection at the segment level can reduce the data size to about 24.14\% of original data, 
which leads to about a 75.86\% reduction. Namely 10TB snapshot data per machine is reduced to 2.414TB.
Similarity-guided local search can further reduce the data size
to about 1.205T, which is  12.05\% of original. Thus it delivers a 50.08\% reduction after the version-based
 deduplication.
The popularity-guided global deduplication with $\sigma=2\% $
reduces the data further to 860GB, namely 8.6\% of the original size. 
So it provides additional 28.63\% reduction. 
The overall memory usage of VC for each physical machine is very small, which is insignificant to other 
hosted cloud services.  In comparison, the 2.4GB per-machine memory usage of SRB is very visible 
to other services.
% even each physical machine in the production environment has 48GB memory.

>>>>

In summary, our experiments show that
segment-level version detection 
reduces the data size to about 24.14\% of original data, 
which is a 75.86\% reduction. Namely 10TB snapshot data per machine is reduced to 2.414TB.
Similarity-guided local search can further reduce the data size
to about 1.205T, which is  12.05\% of original. Thus it delivers a 50.08\% reduction after the version-based
 deduplication.
The popularity-guided global deduplication with $\sigma=2\% $
reduces the data further to 860GB, namely 8.6\% of the original size. 
So it provides additional 28.63\% reduction. 
The overall memory usage of VC for each physical machine is very small, which is insignificant to other 
hosted cloud services.  In comparison, the 2.4GB per-machine memory usage of SRB is very visible 
to other services.
% even each physical machine in the production environment has 48GB memory.
